---
title: "Supervised Learning Methods"
subtitle: "Advanced Modelling 2° Assignment"
author: "Diego Paroli"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook
---

# Project 

...

# Loading libraries

```{r, message=FALSE}
rm(list = ls())
library(tidyverse)
library(httr2)
library(DataExplorer)
library(explore)
library(mice)
```

# Loading data

```{r}
# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_DEMO@DF_DEMO,2.0/A.TL3...LAND_AREA+POP...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/population.csv")
population <- read_csv("data/population.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_DEMO@DF_DEMO,2.0/A.TL3...DEPEND_RATIO+FERT_RATIO...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/demographic.csv")
demographic <- read_csv("data/demographic.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_DEMO@DF_DEMO,2.0/A.TL3...NETMIG+OUTMIG+INMIG+NETMOB+OUTMOB+INMOB...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/migration.csv")
migration <- read_csv("data/migration.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_DEMO@DF_DEMO,2.0/A.TL3...MORT_ICDV_CRUDE_RATIO+MORT_ICDI_CRUDE_RATIO+MORT_ICDJ_CRUDE_RATIO+MORT_INFANT+MORT_STANDARD_RATIO+LFEXP...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/mortality.csv")
mortality <- read_csv("data/mortality.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_SOC@DF_SOC,2.2/A.TL3...MORT_ICDV_CRUDE_RATIO+VEH_THEFT...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/safety.csv")
safety <- read_csv("data/safety.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_SOC@DF_SOC,2.2/A.TL3...VEH_STOCK_ELECTRIC...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/electric_cars.csv")
electric_cars <- read_csv("data/electric_cars.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_SOC@DF_SOC,2.2/A.TL3...BB_ACC...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/internet_access.csv")
internet_access <- read_csv("data/internet_access.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_ECO@DF_ECO,2.0/A.TL3...GDP..Q.USD_PPP_PS?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/gdp.csv")
gdp <- read_csv("data/gdp.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_ECO@DF_ECO,2.0/A.TL3...B5N..Q.USD_PPP_PS?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/income.csv")
income <- read_csv("data/income.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_ECO@DF_ECO,2.0/A.TL3...LAB_PROD._T.Q.USD_PPP_WR?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/labour_productivity.csv")
labour_productivity <- read_csv("data/labour_productivity.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_LABOUR@DF_LAB,2.0/A.TL3...EMP_RATIO.Y15T64._T.PT_POP_SUB?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/employment.csv")
employment <- read_csv("data/employment.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_LABOUR@DF_LAB,2.0/A.TL3...UNE_RATE.Y15T64._T.PT_LF_SUB?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/unemployment.csv")
unemployment <- read_csv("data/unemployment.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_LABOUR@DF_LAB,2.0/A.TL3...LF_RATE_SEXDIF.Y15T64._T.PD?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/gender_diff_labour_force_participation.csv")
gender_diff_labour_force_participation <- read_csv("data/gender_diff_labour_force_participation.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_CLIM@DF_CLIM,1.0/A.TL3...RF_LAND_EXP+CF_LAND_EXP.Y_20...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/floodings.csv")
floodings <- read_csv("data/floodings.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_CLIM@DF_CLIM,1.0/A.TL3...FIRE_LAND_EXP....PT_LAR?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/wildfires.csv")
wildfires <- read_csv("data/wildfires.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_HEALTH@DF_CARE,2.0/A.TL3...HB...10P3HB?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/hospital_beds.csv")
hospital_beds <- read_csv("data/hospital_beds.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_HEALTH@DF_RISK,2.0/A.TL3...PM25_POP_EXP+OBESITY...?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/health_risks.csv")
health_risks <- read_csv("data/health_risks.csv", show_col_types = FALSE)

# data <- request("https://sdmx.oecd.org/public/rest/data/OECD.CFE.EDS,DSD_REG_PAT@DF_PAT,1.0/A.TL3....APPLICATION.PAT.._T.10P6HB?startPeriod=2010&endPeriod=2022&dimensionAtObservation=AllDimensions&format=csvfilewithlabels") |> req_perform()
# writeBin(resp_body_raw(data), "data/patents.csv")
patents <- read_csv("data/patents.csv", show_col_types = FALSE)
```

The data has been obtained through talking directly to the OECD API and the written over .csv files to avoid having to recall the API each time. The code with the queries to the API and the writing of the .csv files have been commented, but they remain available for any future update or for reproducibility. 

# Data cleaning

First cleaning of the variables (selecting the relevant age, sex groups and/or relevant units of measure):

```{r}
# Population and population density
population <- population |> 
  select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
tot_population <- population |> 
  filter(Measure=="Population") |> 
  filter(Age=="Total") |> 
  filter(Sex=="Total") |> 
  filter(`Unit of measure`=="Persons")
pop_density <- population |> 
  filter(Measure=="Population") |> 
  filter(Age=="Total") |> 
  filter(Sex=="Total") |> 
  filter(`Unit of measure`=="Persons per square kilometer")

# Fertility rate and age dependency
demographic <- demographic |> 
  select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
fertility_rate <- demographic |> filter(Measure=="Fertility rate") |> filter(Age == "Total") 
old_dependency_ratio <- demographic |> 
  filter(Measure=="Dependency ratio") |> 
  filter(Age == "65 years or over") |> 
  filter(`Unit of measure` == "Percentage of population aged 15-64 years")

# Net national and international migration
migration <- migration |> 
  select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country) |> 
  filter(Sex == "Total") |> 
  filter(Measure %in% c("Net inter-regional mobility (inflows minus outflows)", "Net international migration (inflows minus outflows)"))
net_total_national_mobility <- migration |> 
  filter(Measure == "Net inter-regional mobility (inflows minus outflows)") |>
  filter(Age == "Total") |> 
  filter(`Unit of measure` == "Percentage of population")
net_total_international_mobility <- migration |> 
  filter(Measure == "Net international migration (inflows minus outflows)")

# Several measure related with mortality
mortality <- mortality |> 
  select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
mortality_rates <- mortality |> 
  filter(Measure == "Age-adjusted mortality rate") |> 
  filter(Sex == "Total")
infant_mortality <- mortality |> 
  filter(Measure == "Infant mortality rate") |> 
  filter(Sex == "Total") 
life_expectancy <- mortality |> 
  filter(Measure == "Life expectancy") |> 
  filter(Sex == "Total") 

# Vehicles thefts
safety <- safety |>
    select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
vehicle_thefts <- safety |> 
  filter(Measure == "Motor vehicle theft") |> 
  filter(`Unit of measure`=="Cases per 100 000 persons") 

# Electric cars
electric_cars <- electric_cars |>
    select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Internet access
internet_access <- internet_access |>
    select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# GDP
gdp <- gdp |>
    select(REF_AREA, `Reference area`, Measure, `Price base`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Primary income
income <- income |>
    select(REF_AREA, `Reference area`, Measure, `Price base`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Labour productivity
labour_productivity <- labour_productivity |>
    select(REF_AREA, `Reference area`, Measure, `Price base`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Employment
employment <- employment |>
    select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Unemployment
unemployment <- unemployment |>
        select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Labour force participation rate, gender difference
gender_diff_labour_force_participation <- gender_diff_labour_force_participation |>
        select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# River flooding, coastal flooding and wildfires
floodings <- floodings |>
        select(REF_AREA, `Reference area`, Measure, `Return period`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
river_flooding <- floodings |>
  filter(Measure == "Land exposure to river flooding")
coastal_flooding <- floodings |>
  filter(Measure == "Land exposure to coastal flooding")

# Wildfires
wildfires <- wildfires |>
        select(REF_AREA, `Reference area`, Measure, `Return period`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Hospital beds
hospital_beds <- hospital_beds |>
        select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)

# Exposure to pollution
pollution <- health_risks |>
        select(REF_AREA, `Reference area`, Measure, Age, Sex, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country) |> 
  filter(Measure == "Population exposure to PM2.5") 

# Patents
patents <- patents |>
  select(REF_AREA, `Reference area`, Measure, `Reference date type`, `Unit of measure`, TIME_PERIOD, OBS_VALUE, COUNTRY, Country)
```

Creating a dataframe storing metadata of all my variables:

```{r}
# List of all my datframes
df_list <- list(
  tot_population, pop_density, 
  mortality_rates, infant_mortality, life_expectancy,
  fertility_rate, old_dependency_ratio,
  net_total_national_mobility, net_total_international_mobility,
  vehicle_thefts, electric_cars, internet_access, 
  gdp, income, labour_productivity, employment, unemployment, 
  gender_diff_labour_force_participation, 
  river_flooding, coastal_flooding, wildfires, 
  hospital_beds, pollution, patents
)

# Function to extract constant columns as a single-row dataframe
get_constant_values <- function(df) {
  constant_cols <- df %>%
    summarise(across(everything(), ~ length(unique(.)) == 1)) %>%
    unlist()
  
  df |> 
    summarise(across(names(df)[constant_cols], first))
}

# Apply the function to all my dataframes and store them in a list 
constant_df <- map(df_list, get_constant_values)

# Turn the list into a dataframe
constant_df <- bind_rows(constant_df)
print(constant_df, n=Inf)
```

Electric vehicles and net international migration need to scaled to the population of a region later on.

Since I have stored all the metadata in a separate dataframe I now delete all the columns referring to metadata and then proceed to take the mean of each of my variables over all the available years: 

```{r}
cols_to_drop <- setdiff(colnames(constant_df), "Measure")

cleaning_function <- function(df) {
  
  measure_name <- unique(df$Measure)
  
  df <- df |> 
    select(-any_of(colnames(constant_df))) |> 
    group_by(REF_AREA, `Reference area`, COUNTRY, Country) |> 
    summarise(!!measure_name := mean(OBS_VALUE, na.rm = TRUE), .groups = "drop")  
}

tot_population <- cleaning_function(tot_population)
pop_density <- cleaning_function(pop_density)
mortality_rates <- cleaning_function(mortality_rates)
infant_mortality <- cleaning_function(infant_mortality)
life_expectancy <- cleaning_function(life_expectancy)
fertility_rate <- cleaning_function(fertility_rate)
old_dependency_ratio <- cleaning_function(old_dependency_ratio)
net_total_national_mobility <- cleaning_function(net_total_national_mobility)
net_total_international_mobility <- cleaning_function(net_total_international_mobility)
vehicle_thefts <- cleaning_function(vehicle_thefts)
electric_cars <- cleaning_function(electric_cars)
internet_access <- cleaning_function(internet_access)
gdp <- cleaning_function(gdp)
income <- cleaning_function(income)
labour_productivity <- cleaning_function(labour_productivity)
employment <- cleaning_function(employment)
unemployment <- cleaning_function(unemployment)
gender_diff_labour_force_participation <- cleaning_function(gender_diff_labour_force_participation)
river_flooding <- cleaning_function(river_flooding)
coastal_flooding <- cleaning_function(coastal_flooding)
wildfires <- cleaning_function(wildfires)
hospital_beds <- cleaning_function(hospital_beds)
pollution <- cleaning_function(pollution)
patents <- cleaning_function(patents)
```

I now merge all datasets together:

```{r}
df_list <- list(
  tot_population, pop_density, 
  mortality_rates, infant_mortality, life_expectancy,
  fertility_rate, old_dependency_ratio,
  net_total_national_mobility, net_total_international_mobility,
  vehicle_thefts, electric_cars, internet_access, 
  gdp, income, labour_productivity, employment, unemployment, 
  gender_diff_labour_force_participation, 
  river_flooding, coastal_flooding, wildfires, 
  hospital_beds, pollution, patents
)
# Merge datasets one by one using full_join
final_data <- df_list[[1]]
# Starting with total population which is the one with the most row allows me to retain all the data below

# Loop through the remaining datasets and join them
for(i in 2:length(df_list)) {
  df_list[[i]] <- df_list[[i]] |> 
    select(-c("Reference area", "COUNTRY", "Country"))
  
  final_data <- full_join(final_data, df_list[[i]], 
                          by = c("REF_AREA"))
}

# View the resulting dataset
str(final_data)
```

Now I have all the datasets together and I will select only EU27's regions

```{r}
# Population.x = tot_population
# Population.y=  pop_density
final_data <- final_data |> 
  rename(tot_population = Population.x,
         pop_density = Population.y,
         region_code = REF_AREA,
         region_name = "Reference area",
         country_code = COUNTRY,
         country_name = Country)

# Deleting extra regions (region code ends with ZZZ)
final_data <- final_data |> 
  filter(!str_ends(region_code, "ZZ"))

eu_codes <- countrycode::codelist |> 
  filter(eu28 == "EU") |> 
  select(iso3c) |> 
  filter(iso3c != "GBR")

eu_data <- final_data |> 
  filter(country_code %in% eu_codes$iso3c)
```

I will now proceed with further cleaning

```{r}
eu_data <- eu_data |> 
  janitor::clean_names() |> 
  rename(
    net_interregional_mobility = net_inter_regional_mobility_inflows_minus_outflows,
    net_international_mobility = net_international_migration_inflows_minus_outflows,
    electric_cars = vehicles_fleet_electric_power_engine,
    share_households_internet = share_of_households_with_internet_broadband_access,
    patents = patent_application_fractionnal_count
  )

eu_data <- eu_data |> 
  mutate(electric_cars = electric_cars/tot_population,
         net_international_mobility = net_international_mobility/tot_population)

eu_data <- eu_data |> 
  select(-tot_population)
```

Creating a factor variable which will be our target variable for the classification methods:

```{r}
eu_data <- eu_data |> 
  mutate(region = as.factor(case_when(
    country_code %in% c("AUT", "BEL", "DEU", "FRA", "LUX", "NLD") ~ "Western Europe",
    country_code %in% c("GRC", "ESP", "ITA", "MLT", "PRT", "IRL") ~ "Southern Europe",
    country_code %in% c("DNK", "EST", "FIN", "LTU", "LVA", "SWE") ~ "Northern Europe",
    country_code %in% c("CZE", "HRV", "HUN", "POL", "ROU", "SVK", "SVN", "BGR") ~ "Eastern Europe")))
```


# Exploratory Data Analysis

```{r}
# No data on cyprus
unique(eu_data$country_name)
```



```{r}
explore_tbl(eu_data)
plot_intro(eu_data)
plot_missing(eu_data)
```

There are some variables which have worrying numbers of missing values.

I am going to delete `net_international_mobility` because there is already a closely related (`net_interrregional_mobility`). I will also delete `primary_incomes_net` as that is likely to be related with GDP, and `share_households_internet` so as to retain only variables that have at least <50% of NAs

```{r}
eu_data <- eu_data |>
  select(-c(net_international_mobility, primary_incomes_net, share_households_internet))
```

```{r}
eu_data |> 
  mutate(na = rowSums(is.na(eu_data))) |> 
  group_by(country_code) |>  
  summarize(mean_na = round(mean(na, na.rm = TRUE), 2)) |> 
  print(n=Inf)
```

We can see that Eastern Europe's states and small states tend to have a higher number of missing values on average.

```{r}
summary(eu_data)
```

```{r, fig.height=21}
explore_all(eu_data %>% select(where(is.numeric)))
```

Class distribution of the classification target variable:

```{r}
eu_data |>  count(region)
```

Conditional boxplots of my predictors based on the classification target variable

```{r, warning=FALSE, fig.height=11}
plot_boxplot(eu_data, by="region", ncol = 2, nrow = 5)
```

Correlation of numerical variables. By looking at the row of `gross_domestic_product`

```{r, fig.height=8}
plot_correlation(eu_data, type="continuous", cor_args=list("use" = "pairwise.complete.obs"))
```


# Imputation of missing values 

```{r}

```


# 

```{r}

```

